---
title: "Reproduce Secondary Outcomes - HIV Self-Testing Study"
format: 
  html:
    toc: true
    toc-depth: 3
    code-fold: false
    embed-resources: true
execute:
  warning: false
  message: false
---

This document reproduces the Secondary Outcomes analyses from Multimedia Appendix 3.

We will reproduce the following key findings:

1. **Q15_5**: "People in my life would leave if I had HIV" - Fisher's exact test
2. **Q94_13**: "I think that new HIV/AIDS treatments can eradicate the virus from your body" - Wilcoxon rank test
3. **Q14_3**: "I could not be friends with someone who has HIV/AIDS" - Wilcoxon rank test

```{r}
#| label: load-libraries

library(tidyverse)
library(knitr)
library(kableExtra)
```

## Data Preparation

```{r}
#| label: read-data

# Read the raw data
raw_data <- read.csv("data/CTN_FINAL.csv", stringsAsFactors = FALSE)

# Check total observations
cat("Total observations in raw data:", nrow(raw_data), "\n")

# Filter to per-protocol sample
df <- raw_data %>%
  filter(PO_FLAG == "Include", WAVE != "3")

cat("Per-protocol sample size:", nrow(df), "\n")
```

## Create Analysis Variables

```{r}
#| label: create-variables

analysis_df <- df %>%
  mutate(
    ordered_kit = case_when(
      ORA_REDEEMED == "Yes" ~ "Ordered",
      ORA_REDEEMED %in% c("No", "Over 60 days") ~ "Did not order",
      TRUE ~ NA_character_
    )
  )

# Check the distribution
cat("\n=== Distribution of Test Kit Orders ===\n")
table(analysis_df$ordered_kit, useNA = "ifany")
```

## Analysis 1: Q15_5 - "People in my life would leave if I had HIV"

```{r}
#| label: analysis-q15-5

# Q15_5: People in my life would leave if I had HIV
# 1 = Agree, 2 = Disagree

# Create the response variable
analysis_df <- analysis_df %>%
  mutate(
    Q15_5_response = case_when(
      Q15_5 == 1 ~ "Agree",
      Q15_5 == 2 ~ "Disagree",
      TRUE ~ NA_character_
    )
  )

# Filter to valid responses for this question
q15_5_data <- analysis_df %>%
  filter(!is.na(ordered_kit) & !is.na(Q15_5_response))


# Cross-tabulation
cross_tab <- table(q15_5_data$ordered_kit, q15_5_data$Q15_5_response)
print(cross_tab)

# Calculate percentages within each group
cat("\n--- Percentages by group ---\n")
ordered_data <- q15_5_data %>% filter(ordered_kit == "Ordered")
not_ordered_data <- q15_5_data %>% filter(ordered_kit == "Did not order")

n_ordered <- nrow(ordered_data)
n_not_ordered <- nrow(not_ordered_data)

cat("Ordered test kit (n=", n_ordered, "):\n", sep = "")
cat("  Agree: ", sum(ordered_data$Q15_5_response == "Agree"), "/", n_ordered, 
    " (", round(sum(ordered_data$Q15_5_response == "Agree")/n_ordered*100, 1), "%)\n", sep = "")
cat("  Disagree: ", sum(ordered_data$Q15_5_response == "Disagree"), "/", n_ordered, 
    " (", round(sum(ordered_data$Q15_5_response == "Disagree")/n_ordered*100, 1), "%)\n", sep = "")

cat("\nDid not order test kit (n=", n_not_ordered, "):\n", sep = "")
cat("  Agree: ", sum(not_ordered_data$Q15_5_response == "Agree"), "/", n_not_ordered, 
    " (", round(sum(not_ordered_data$Q15_5_response == "Agree")/n_not_ordered*100, 1), "%)\n", sep = "")
cat("  Disagree: ", sum(not_ordered_data$Q15_5_response == "Disagree"), "/", n_not_ordered, 
    " (", round(sum(not_ordered_data$Q15_5_response == "Disagree")/n_not_ordered*100, 1), "%)\n", sep = "")

# Fisher's Exact Test
cat("\n--- Fisher's Exact Test ---\n")
fisher_result_q15_5 <- fisher.test(cross_tab)
print(fisher_result_q15_5)

# Create summary table
q15_5_summary <- tibble(
  Statement = "People in my life would leave if I had HIV",
  Response = c("Agree", "Disagree"),
  `Ordered test kit n` = c(sum(ordered_data$Q15_5_response == "Agree"),
                           sum(ordered_data$Q15_5_response == "Disagree")),
  `Ordered test kit N` = n_ordered,
  `Ordered test kit %` = c(round(sum(ordered_data$Q15_5_response == "Agree")/n_ordered*100, 1),
                           round(sum(ordered_data$Q15_5_response == "Disagree")/n_ordered*100, 1)),
  `Did not order n` = c(sum(not_ordered_data$Q15_5_response == "Agree"),
                        sum(not_ordered_data$Q15_5_response == "Disagree")),
  `Did not order N` = n_not_ordered,
  `Did not order %` = c(round(sum(not_ordered_data$Q15_5_response == "Agree")/n_not_ordered*100, 1),
                        round(sum(not_ordered_data$Q15_5_response == "Disagree")/n_not_ordered*100, 1)),
  `P-value` = c(round(fisher_result_q15_5$p.value, 3), "—")
)

kable(q15_5_summary, caption = "Table (c) - Attitudes toward HIV testing: Q15_5") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Analysis 2: Q94_13 - "I think that new HIV/AIDS treatments can eradicate the virus from your body"

```{r}
#| label: analysis-q94-13

# Q94_13: I think that new HIV/AIDS treatments can eradicate the virus from your body
# Scale: 1 (Strongly disagree) to 7 (Strongly agree)

# Filter to valid responses
q94_13_data <- analysis_df %>%
  filter(!is.na(ordered_kit) & !is.na(Q94_13) & Q94_13 != 9999)

# Summary statistics by group
ordered_q94 <- q94_13_data %>% filter(ordered_kit == "Ordered")
not_ordered_q94 <- q94_13_data %>% filter(ordered_kit == "Did not order")

cat("Ordered test kit (n=", nrow(ordered_q94), "):\n", sep = "")
cat("  Mean (SD):", round(mean(ordered_q94$Q94_13), 1), "(", round(sd(ordered_q94$Q94_13), 1), ")\n")
cat("  Median (IQR):", median(ordered_q94$Q94_13), "(", quantile(ordered_q94$Q94_13, 0.25), "-", quantile(ordered_q94$Q94_13, 0.75), ")\n")

cat("\nDid not order test kit (n=", nrow(not_ordered_q94), "):\n", sep = "")
cat("  Mean (SD):", round(mean(not_ordered_q94$Q94_13), 1), "(", round(sd(not_ordered_q94$Q94_13), 1), ")\n")
cat("  Median (IQR):", median(not_ordered_q94$Q94_13), "(", quantile(not_ordered_q94$Q94_13, 0.25), "-", quantile(not_ordered_q94$Q94_13, 0.75), ")\n")

# Wilcoxon Rank Sum Test
cat("\n--- Wilcoxon Rank Sum Test ---\n")
wilcox_result_q94_13 <- wilcox.test(ordered_q94$Q94_13, not_ordered_q94$Q94_13)
print(wilcox_result_q94_13)

# Create summary table
q94_13_summary <- tibble(
  Statement = "I think that new HIV/AIDS treatments can eradicate the virus from your body",
  `Ordered test kit (n)` = nrow(ordered_q94),
  `Ordered Mean (SD)` = paste0(round(mean(ordered_q94$Q94_13), 1), " (", round(sd(ordered_q94$Q94_13), 1), ")"),
  `Did not order (n)` = nrow(not_ordered_q94),
  `Did not order Mean (SD)` = paste0(round(mean(not_ordered_q94$Q94_13), 1), " (", round(sd(not_ordered_q94$Q94_13), 1), ")"),
  `P-value (Wilcoxon)` = round(wilcox_result_q94_13$p.value, 3)
)

kable(q94_13_summary, caption = "Table (d) - Attitudes toward HIV treatment: Q94_13") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Analysis 3: Q14_3 - "I could not be friends with someone who has HIV/AIDS"

```{r}
#| label: analysis-q14-3

# Q14_3: I could not be friends with someone who has HIV/AIDS
# Scale: 1 = Strongly agree, 2 = Agree, 3 = Somewhat agree, 4 = Neither agree nor disagree,
#        5 = Somewhat disagree, 6 = Disagree, 7 = Strongly disagree

# Filter to valid responses
q14_3_data <- analysis_df %>%
  filter(!is.na(ordered_kit) & !is.na(Q14_3) & Q14_3 != 9999)

# Summary statistics by group
ordered_q14 <- q14_3_data %>% filter(ordered_kit == "Ordered")
not_ordered_q14 <- q14_3_data %>% filter(ordered_kit == "Did not order")

cat("Ordered test kit (n=", nrow(ordered_q14), "):\n", sep = "")

print(table(ordered_q14$Q14_3))

cat("\nDid not order test kit (n=", nrow(not_ordered_q14), "):\n", sep = "")

print(table(not_ordered_q14$Q14_3))

# Wilcoxon Rank Sum Test
cat("\n--- Wilcoxon Rank Sum Test ---\n")
wilcox_result_q14_3 <- wilcox.test(ordered_q14$Q14_3, not_ordered_q14$Q14_3)
print(wilcox_result_q14_3)

# Create detailed frequency table
response_labels <- c("1" = "Strongly agree", "2" = "Agree", "3" = "Somewhat agree", 
                     "4" = "Neither agree nor disagree", "5" = "Somewhat disagree", 
                     "6" = "Disagree", "7" = "Strongly disagree")

ordered_freq <- table(ordered_q14$Q14_3)
not_ordered_freq <- table(not_ordered_q14$Q14_3)

# Create full frequency table including all response levels
all_levels <- as.character(1:7)
ordered_counts <- sapply(all_levels, function(x) ifelse(x %in% names(ordered_freq), ordered_freq[x], 0))
not_ordered_counts <- sapply(all_levels, function(x) ifelse(x %in% names(not_ordered_freq), not_ordered_freq[x], 0))

q14_3_detail <- tibble(
  `Response` = response_labels[all_levels],
  `Ordered n` = ordered_counts,
  `Ordered N` = nrow(ordered_q14),
  `Ordered %` = round(ordered_counts / nrow(ordered_q14) * 100, 1),
  `Not ordered n` = not_ordered_counts,
  `Not ordered N` = nrow(not_ordered_q14),
  `Not ordered %` = round(not_ordered_counts / nrow(not_ordered_q14) * 100, 1)
)

kable(q14_3_detail, caption = "Table (e) - HIV-related stigma: Q14_3 - I could not be friends with someone who has HIV/AIDS") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  footnote(general = paste0("P-value (Wilcoxon rank test) = ", round(wilcox_result_q14_3$p.value, 3)))
```

# Conclusions

The secondary outcomes have been reproduced. All p-values match the manuscript (the result of the Analysis 3 is slightly different from the manuscript with a p-value of 0.033, but the conclusion is the same).